---
title: "GL_station2route"
author: "Huajie Yang"
date: "11/7/2019"
output: html_document
---

# Setting 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
options(scipen=100)
options(digits=3)

if (!require("pacman")) {install.packages("pacman"); library(pacman)}
p_load(dplyr, lubridate, ggplot2, stargazer, stringr)
```


# Load original data 
```{r}
# Load data 
load("~/OrangeGreenLines/output/intermediate/data_5min_0506_green.RData")
load("~/OrangeGreenLines/output/intermediate/data_5min_0607_green.RData")

data_5min_0506_green <- data_5min_0506_green %>%
                        filter(starttime!="2006-09-12 00:00:00-07")

data_5min_0607_green <- data_5min_0607_green %>%
                        filter(starttime!="2007-09-12 00:00:00-07") %>%
                        filter(!(detector_id %in% c(1949, 1950, 1951, 1953, 1954, 1955)))

data_5min_0507_green <- rbind(data_5min_0506_green, data_5min_0607_green) %>%
                        mutate(date_time=ymd_hms(substr(starttime, 1, nchar(starttime)-3)),
                               date=as.Date(starttime),
                               day_week=weekdays(date),
                               before_after=ifelse(date < ymd("2009-09-12"), 0, 1),
                               time_hour=hour(date_time),
                               time_min5=minute(date_time),
                               AM_PM=ifelse(time_hour %in% c(6, 7, 8, 9), "AM", "NonPeak"),
                               AM_PM=ifelse(time_hour %in% c(16, 17, 18, 19), "PM", AM_PM),
                               ec_group=ifelse(highwayid %in% c(1,2), "control", "experimental")) %>%
                        select(highwayid, stationid, lanenumber, detector_id, starttime, date_time, date, day_week, time_hour, time_min5,
                               AM_PM, speed)  %>% #
                        arrange(highwayid, stationid, lanenumber, detector_id, date_time)

# I-84 east bound stations
# 1097 is too close to downtown
# , locationtext, length, upstream,
# downstream, numberlanes, length_mid, downstream_mile,
# upstream_mile, opposite_stationid, start_date, end_date
stations_meta <- read.csv("data/metadata/stations_metadata.csv", stringsAsFactors = F) # station metadata


stations_I84 <- stations_meta %>%
  filter(highwayid %in% c(7, 8), start_date=="2004-01-01 00:00:00-08")

# East bound
stations_I84_7 <- stations_I84 %>%
                  filter(highwayid %in% c(7), stationid %in% c(1055, 1056, 1057, 1127)) %>%
                  select(stationid, highwayid, milepost)

data_5min_0507_green_I84_east <- data_5min_0507_green %>%
                                  filter(stationid %in% c(1055, 1056, 1057, 1127))  %>%
                                  filter(lanenumber==1, AM_PM %in% c("AM", "PM")) %>%
                                  # select(-date_time) %>%
                                  left_join(stations_I84_7) %>%
                                  arrange(date, time_hour, time_min5, stationid)%>%
                                  mutate(id=paste(highwayid, date, time_hour, time_min5, sep="_"))


save(data_5min_0507_green_I84_east, file="~/OrangeGreenLines/output/intermediate/data_5min_0507_green_I84_east.RData")

load("~/OrangeGreenLines/output/intermediate/data_5min_0507_green_I84_east.RData")
```

# Check missing values 
```{r}
test <- data_5min_0507_green_I84_east %>%
         group_by(highwayid, date, time_hour, time_min5) %>%
         summarise(freq=n())
table(test$freq)
```

# reorganize 
```{r}
# 730 is the number of days: length(seq(as.Date("2005-09-12"), as.Date("2007-09-11"), by="days")) 
# 8: 8 hours: c(6:9, 16:19)
# 12: time intervals: seq(0, 55, 5)
# 4 stations: c(1055, 1056, 1057, 1127)
full_df_I84_east <- data.frame(date=rep(seq(as.Date("2005-09-12"), as.Date("2007-09-11"), by="days"), each=8*12*4),
                               time_hour=rep(c(6:9, 16:19), each=12*4, times=730),
                               time_min5=rep(seq(0, 55, 5), each=4, times=730*8),
                               stationid=rep(c(1055, 1056, 1057, 1127), 730*8*12))  %>%
                    left_join(data_5min_0507_green_I84_east) %>%
                    mutate(time_id=paste(date, time_hour, time_min5, sep="_"),
                           AM_PM=ifelse(time_hour %in% c(6:9), "AM", "PM"))

# Check whether all observations are missing at time points 
speed_NA_df <- full_df_I84_east %>%
                group_by(date, time_hour, time_min5) %>%
                summarise(speed_NA=all(is.na(speed)))  %>%
                filter(speed_NA==TRUE) %>%
                mutate(time_id=paste(date, time_hour, time_min5, sep="_"))
# Calculate peak periods average speed of previous/later day 
full_df_I84_east_day_pp <- full_df_I84_east %>%
                        select(stationid, date, day_week, time_hour, time_min5, starttime, AM_PM, speed) %>%
                        mutate(week_num=week(date),
                               day_week=weekdays(date),
                               year=year(date))  %>%
                        filter(!(day_week %in% c("Saturday", "Sunday")))  %>%
                        group_by(stationid, year, week_num, date, AM_PM) %>%
                        summarise(speed_avg=mean(speed, na.rm=TRUE),
                                  speed_median=median(speed, na.rm=TRUE)) %>%
                        ungroup() %>%
                        group_by(stationid, year, week_num, AM_PM) %>%
                        mutate(speed_avg_lag1=lag(speed_avg),
                               speed_avg_lag2=lag(speed_avg, 2),
                               speed_avg_lag3=lag(speed_avg, 3),
                               speed_avg_lag4=lag(speed_avg, 4),
                               
                               speed_avg_lead1=lead(speed_avg),
                               speed_avg_lead2=lead(speed_avg, 2),
                               speed_avg_lead3=lead(speed_avg, 3),
                               speed_avg_lead4=lead(speed_avg, 4)
                        ) %>%
                        ungroup() %>%
                        mutate(day_week=weekdays(date))  %>%
                        select(stationid, date, week_num, AM_PM, day_week, year, speed_avg, speed_median,
                               speed_avg_lag1, speed_avg_lag2, speed_avg_lag3, speed_avg_lag4, 
                               speed_avg_lead1, speed_avg_lead2, speed_avg_lead3, speed_avg_lead4) %>%
                        arrange(stationid, year, week_num, AM_PM) %>%
                        # mutate(stationid=as.factor(stationid)) %>%
                        as.data.frame()

# get peak periods average speed the same day of week of last/previous week 
full_df_I84_east_week_pp <- full_df_I84_east_day_pp %>%
                            arrange(stationid, AM_PM, date)  %>%
                            group_by(stationid, AM_PM)  %>%
                            mutate(speed_avg_lag5=lag(speed_avg, 5),
                                   speed_avg_lag10=lag(speed_avg, 10),
                                   speed_avg_lead5=lead(speed_avg, 5),
                                   speed_avg_lead10=lead(speed_avg, 10)) %>%
                            select(stationid, date, week_num, AM_PM, day_week, year, speed_avg, 
                                   speed_avg_lag5, speed_avg_lag10, speed_avg_lead5, speed_avg_lead10) %>%
                            as.data.frame()

# Get eariler or later observations 
full_df_I84_east_time <- full_df_I84_east %>%
                          select(stationid, date, time_hour, time_min5, starttime, AM_PM, speed) %>%
                          arrange(stationid, date, time_hour, time_min5)  %>%
                          group_by(stationid, date, AM_PM)  %>%
                          mutate(speed_tlag1=lag(speed),
                                 speed_tlag2=lag(speed, 2),
                                 speed_tlag3=lag(speed, 3),
                                 speed_tlag4=lag(speed, 4),
                                 speed_tlag5=lag(speed, 5),
                                 speed_tlag6=lag(speed, 6),
                                 speed_tlag7=lag(speed, 7),
                                 speed_tlag8=lag(speed, 8),
                                 speed_tlag9=lag(speed, 9),
                                 speed_tlag10=lag(speed, 10),
                                 speed_tlag11=lag(speed, 11),
                                 speed_tlag12=lag(speed, 12),
                                 
                                 speed_tlag13=lag(speed, 13),
                                 speed_tlag14=lag(speed, 14),
                                 speed_tlag15=lag(speed, 15),
                                 speed_tlag16=lag(speed, 16),
                                 speed_tlag17=lag(speed, 17),
                                 speed_tlag18=lag(speed, 18),
                                 speed_tlag19=lag(speed, 19),
                                 speed_tlag20=lag(speed, 20),
                                 speed_tlag21=lag(speed, 21),
                                 speed_tlag22=lag(speed, 22),
                                 speed_tlag23=lag(speed, 23),
                                 speed_tlag24=lag(speed, 24),
                                 
                                 speed_tlag25=lag(speed, 25),
                                 speed_tlag26=lag(speed, 26),
                                 speed_tlag27=lag(speed, 27),
                                 speed_tlag28=lag(speed, 28),
                                 speed_tlag29=lag(speed, 29),
                                 speed_tlag30=lag(speed, 30),
                                 speed_tlag31=lag(speed, 31),
                                 speed_tlag32=lag(speed, 32),
                                 speed_tlag33=lag(speed, 33),
                                 speed_tlag34=lag(speed, 34),
                                 speed_tlag35=lag(speed, 35),
                                 speed_tlag36=lag(speed, 36),
                                 
                                 speed_tlag37=lag(speed, 37),
                                 speed_tlag38=lag(speed, 38),
                                 speed_tlag39=lag(speed, 39),
                                 speed_tlag40=lag(speed, 40),
                                 speed_tlag41=lag(speed, 41),
                                 speed_tlag42=lag(speed, 42),
                                 speed_tlag43=lag(speed, 43),
                                 speed_tlag44=lag(speed, 44),
                                 speed_tlag45=lag(speed, 45),
                                 speed_tlag46=lag(speed, 46),
                                 speed_tlag47=lag(speed, 47),
                                 
                                 speed_tlead1=lead(speed),
                                 speed_tlead2=lead(speed, 2),
                                 speed_tlead3=lead(speed, 3),
                                 speed_tlead4=lead(speed, 4),
                                 speed_tlead5=lead(speed, 5),
                                 speed_tlead6=lead(speed, 6),
                                 speed_tlead7=lead(speed, 7),
                                 speed_tlead8=lead(speed, 8),
                                 speed_tlead9=lead(speed, 9),
                                 speed_tlead10=lead(speed, 10),
                                 speed_tlead11=lead(speed, 11),
                                 speed_tlead12=lead(speed, 12),
                                 
                                 speed_tlead13=lead(speed, 13),
                                 speed_tlead14=lead(speed, 14),
                                 speed_tlead15=lead(speed, 15),
                                 speed_tlead16=lead(speed, 16),
                                 speed_tlead17=lead(speed, 17),
                                 speed_tlead18=lead(speed, 18),
                                 speed_tlead19=lead(speed, 19),
                                 speed_tlead20=lead(speed, 20),
                                 speed_tlead21=lead(speed, 21),
                                 speed_tlead22=lead(speed, 22),
                                 speed_tlead23=lead(speed, 23),
                                 speed_tlead24=lead(speed, 24),
                                 
                                 speed_tlead25=lead(speed, 25),
                                 speed_tlead26=lead(speed, 26),
                                 speed_tlead27=lead(speed, 27),
                                 speed_tlead28=lead(speed, 28),
                                 speed_tlead29=lead(speed, 29),
                                 speed_tlead30=lead(speed, 30),
                                 speed_tlead31=lead(speed, 31),
                                 speed_tlead32=lead(speed, 32),
                                 speed_tlead33=lead(speed, 33),
                                 speed_tlead34=lead(speed, 34),
                                 speed_tlead35=lead(speed, 35),
                                 speed_tlead36=lead(speed, 36),
                                 
                                 speed_tlead37=lead(speed, 37),
                                 speed_tlead38=lead(speed, 38),
                                 speed_tlead39=lead(speed, 39),
                                 speed_tlead40=lead(speed, 40),
                                 speed_tlead41=lead(speed, 41),
                                 speed_tlead42=lead(speed, 42),
                                 speed_tlead43=lead(speed, 43),
                                 speed_tlead44=lead(speed, 44),
                                 speed_tlead45=lead(speed, 45),
                                 speed_tlead46=lead(speed, 46),
                                 speed_tlead47=lead(speed, 47)
                                 )  %>%
                          as.data.frame() %>%
                          mutate(time_id=paste(date, time_hour, time_min5, sep="_")) %>%
                          left_join(full_df_I84_east_day_pp) %>%
                          left_join(full_df_I84_east_week_pp) %>%
                          arrange(stationid, date, time_hour, time_min5)

full_df_I84_east_station <- full_df_I84_east %>%
                            select(stationid, date, time_hour, time_min5, starttime, AM_PM, speed) %>%
                            arrange(date, time_hour, time_min5, stationid)  %>%
                            group_by(date, AM_PM)  %>%
                            mutate(speed_slag1=lag(speed),
                                   speed_slag2=lag(speed, 5),
                                   speed_slag3=lag(speed, 9),
                                   speed_slead1=lead(speed),
                                   speed_slead2=lead(speed, 5),
                                   speed_slead3=lead(speed, 9)) %>%
                            mutate(time_id=paste(date, time_hour, time_min5, sep="_"),
                                   speed_slag1=ifelse(stationid==1055, NA, speed_slag1),
                                   speed_slag2=ifelse(stationid==1055, NA, speed_slag2),
                                   speed_slag3=ifelse(stationid==1055, NA, speed_slag3),
                                   speed_slead1=ifelse(stationid==1127, NA, speed_slead1),
                                   speed_slead2=ifelse(stationid==1127, NA, speed_slead2),
                                   speed_slead3=ifelse(stationid==1127, NA, speed_slead3)
                                   ) %>%
                          as.data.frame()
```

# Explore inputing strategy 

                                         

## Plot 
```{r}
# Plot 5 minute interval data by stationid
p_5min_sid <- ggplot(data=data_5min_0507_green_I84_east %>% mutate(stationid=as.factor(stationid)), aes(x=date_time, y=speed, group=stationid)) +
  geom_point(aes(colour=stationid), size=0.25) + 
  #geom_smooth(aes(linetype=stationid), colour="yellow", size=0.75) + 
  ggtitle("5-min speed by stationid over time") + #+  ylim(0, 50)
  labs( x="Date", y="speed (mph)") +  facet_grid( ~ AM_PM) + 
  theme(plot.title = element_text(hjust = 0.5))

p_5min_sid

# plot daily average speed during AM/PM by stationid 
p_day_sid <- ggplot(data=full_df_I84_east_day_pp %>% mutate(stationid=as.factor(stationid)), aes(x=date, y=speed_avg, group=stationid)) +
  geom_point(aes(colour=stationid), size=0.5) + 
  #geom_smooth(aes(linetype=stationid), colour="yellow", size=0.75) + 
  ggtitle("Daily AM/PM speed by stationid over time") + #+  ylim(0, 50)
  labs( x="Date", y="speed (mph)") +  facet_grid( ~ AM_PM) + 
  theme(plot.title = element_text(hjust = 0.5))

p_day_sid
```

## Correlation 
```{r}
cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_avg", "speed_median",
                              "speed_tlag1", "speed_tlag2", "speed_tlag3", 
                              "speed_tlag4", "speed_tlag5", "speed_tlag6",
                              "speed_tlag7", "speed_tlag8", "speed_tlag9",
                              "speed_tlag10", "speed_tlag11", "speed_tlag12")], 
    use = "complete.obs")

cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_tlag13", "speed_tlag14", "speed_tlag15", 
                              "speed_tlag16", "speed_tlag17", "speed_tlag18",
                              "speed_tlag19", "speed_tlag20", "speed_tlag21",
                              "speed_tlag22", "speed_tlag23", "speed_tlag24")], 
    use = "complete.obs")

cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_tlag25", "speed_tlag26", "speed_tlag27", 
                              "speed_tlag28", "speed_tlag29", "speed_tlag30",
                              "speed_tlag31", "speed_tlag32", "speed_tlag33",
                              "speed_tlag34", "speed_tlag35", "speed_tlag36")], 
    use = "complete.obs")


cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_tlag37", "speed_tlag38", "speed_tlag39", 
                              "speed_tlag40", "speed_tlag41", "speed_tlag42",
                              "speed_tlag43", "speed_tlag44", "speed_tlag45",
                              "speed_tlag46", "speed_tlag47")], 
    use = "complete.obs")

cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_avg_lag1", "speed_avg_lag2", "speed_avg_lag3", "speed_avg_lag4")], 
    use = "complete.obs")

cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_avg_lead1", "speed_avg_lead2", "speed_avg_lead3", "speed_avg_lead4")], 
    use = "complete.obs")

cor(full_df_I84_east_time$speed, 
    full_df_I84_east_time[, c("speed_avg_lag5", "speed_avg_lag10", "speed_avg_lead5", "speed_avg_lead10")], 
    use = "complete.obs")

cor(full_df_I84_east_station$speed, 
    full_df_I84_east_station[, c("speed_slead1", "speed_slead2", "speed_slead3")], 
    use = "complete.obs")
```

# Inpute the the missing values with lter/earlier observation 

## 15 min range 
```{r}
full_df_I84_east_time_15min <- full_df_I84_east_time %>%
                         filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
                         select(-starttime, -time_id) %>%
                         mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, speed_tlead1),
                                speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
                                speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
                                speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
                                speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new)
                                )

# 93.3 percent missing values are not inputed 
nrow(full_df_I84_east_time_15min %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_15min)
```

## 30 min range 
```{r}
full_df_I84_east_time_sub_30 <- full_df_I84_east_time %>%
  filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
  select(-starttime, -time_id) %>%
  mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new))

# 90.4 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_30 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_30)

```

## An hour range 
```{r}
full_df_I84_east_time_sub_60 <- full_df_I84_east_time %>%
  filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
  select(-starttime, -time_id) %>%
  mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag10), speed_tlag10, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead10), speed_tlead10, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag11), speed_tlag11, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead11), speed_tlead11, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag12), speed_tlag12, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead12), speed_tlead12, speed_new)
         )

# 86.9 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_60 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_60)
```

## Two hours range 
```{r}
full_df_I84_east_time_sub_120 <- full_df_I84_east_time %>%
  filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
  select(-starttime, -time_id) %>%
  mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag10), speed_tlag10, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead10), speed_tlead10, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag11), speed_tlag11, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead11), speed_tlead11, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag12), speed_tlag12, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead12), speed_tlead12, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag13), speed_tlag13, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead13), speed_tlead13, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag14), speed_tlag14, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead14), speed_tlead14, speed_new),

         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag15), speed_tlag15, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead15), speed_tlead15, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag16), speed_tlag16, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead16), speed_tlead16, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag17), speed_tlag17, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead17), speed_tlead17, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag18), speed_tlag18, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead18), speed_tlead18, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag19), speed_tlag19, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead19), speed_tlead19, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag20), speed_tlag20, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead20), speed_tlead20, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag21), speed_tlag21, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead21), speed_tlead21, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag22), speed_tlag22, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead22), speed_tlead22, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag23), speed_tlag23, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead23), speed_tlead23, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag24), speed_tlag24, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead24), speed_tlead24, speed_new)
         )

# 82.9 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_120 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_120)
```

## 4 hours
```{r}
# 4 hours range 
full_df_I84_east_time_sub_240 <- full_df_I84_east_time %>%
  filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
  select(-starttime, -time_id) %>%
  mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag10), speed_tlag10, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead10), speed_tlead10, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag11), speed_tlag11, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead11), speed_tlead11, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag12), speed_tlag12, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead12), speed_tlead12, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag13), speed_tlag13, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead13), speed_tlead13, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag14), speed_tlag14, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead14), speed_tlead14, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag15), speed_tlag15, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead15), speed_tlead15, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag16), speed_tlag16, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead16), speed_tlead16, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag17), speed_tlag17, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead17), speed_tlead17, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag18), speed_tlag18, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead18), speed_tlead18, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag19), speed_tlag19, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead19), speed_tlead19, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag20), speed_tlag20, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead20), speed_tlead20, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag21), speed_tlag21, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead21), speed_tlead21, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag22), speed_tlag22, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead22), speed_tlead22, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag23), speed_tlag23, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead23), speed_tlead23, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag24), speed_tlag24, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead24), speed_tlead24, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag25), speed_tlag25, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead25), speed_tlead25, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag26), speed_tlag26, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead26), speed_tlead26, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag27), speed_tlag27, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead27), speed_tlead27, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag28), speed_tlag28, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead28), speed_tlead28, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag29), speed_tlag29, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead29), speed_tlead29, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag30), speed_tlag30, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead30), speed_tlead30, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag31), speed_tlag31, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead31), speed_tlead31, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag32), speed_tlag32, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead32), speed_tlead32, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag33), speed_tlag33, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead33), speed_tlead33, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag34), speed_tlag34, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead34), speed_tlead34, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag35), speed_tlag35, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead35), speed_tlead35, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag36), speed_tlag36, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead36), speed_tlead36, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag37), speed_tlag37, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead37), speed_tlead37, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag38), speed_tlag38, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead38), speed_tlead38, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag39), speed_tlag39, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead39), speed_tlead39, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag40), speed_tlag40, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead40), speed_tlead40, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag41), speed_tlag41, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead41), speed_tlead41, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag42), speed_tlag42, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead42), speed_tlead42, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag43), speed_tlag43, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead43), speed_tlead43, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag44), speed_tlag44, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead44), speed_tlead44, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag45), speed_tlag45, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead45), speed_tlead45, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag46), speed_tlag46, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead46), speed_tlead46, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag47), speed_tlag47, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead47), speed_tlead47, speed_new),
         
  )

# 80.3 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_240 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_240)

```

## 90 min range & previous/later day: time lag 1-9, speed_avg, spee_avg lag 1:4
```{r}
full_df_I84_east_time_sub_day_PP_90 <- full_df_I84_east_time %>%
                                    filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
                                    select(-starttime, -time_id) %>%
                                    mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg), speed_avg, speed_new), # average speed
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag1), speed_avg_lag1, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead1), speed_avg_lead1, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag2), speed_avg_lag2, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead2), speed_avg_lead2, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag3), speed_avg_lag3, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead3), speed_avg_lead3, speed_new),
                                           
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag4), speed_avg_lag4, speed_new),
                                           speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead4), speed_avg_lead4, speed_new)
                                    )

# 66.8 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_day_PP_90 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_day_PP_90)
```

## 90 minute range & previous/later days & the same day of week of previous or next week 
```{r}
# Inputing order: time lag 1-9, speed_avg, spee_avg lag 1:4, the same day of week of last/next week
full_df_I84_east_time_sub_week_day_PP_90 <- full_df_I84_east_time %>%
  filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
  select(-starttime, -time_id) %>%
  mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg), speed_avg, speed_new), # average speed
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag1), speed_avg_lag1, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead1), speed_avg_lead1, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag2), speed_avg_lag2, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead2), speed_avg_lead2, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag3), speed_avg_lag3, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead3), speed_avg_lead3, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag4), speed_avg_lag4, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead4), speed_avg_lead4, speed_new),
         
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag5), speed_avg_lag5, speed_new),
         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead5), speed_avg_lead5, speed_new)
  )

# 59.6 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_week_day_PP_90 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_week_day_PP_90)
```

## 90 minute range & previous/later days & the same day of week of previous or next 2 week 
```{r}
full_df_I84_east_time_sub_2week_day_PP_90 <- full_df_I84_east_time %>%
                                  filter(is.na(speed), !(time_id %in% speed_NA_df$time_id))  %>%
                                  select(-starttime, -time_id) %>%
                                  mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg), speed_avg, speed_new), # average speed
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag1), speed_avg_lag1, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead1), speed_avg_lead1, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag2), speed_avg_lag2, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead2), speed_avg_lead2, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag3), speed_avg_lag3, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead3), speed_avg_lead3, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag4), speed_avg_lag4, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead4), speed_avg_lead4, speed_new),
                                         
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag10), speed_avg_lag10, speed_new),
                                         speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead10), speed_avg_lead10, speed_new)
                                  )

# 53.9 percent missing values are not inputed 
nrow(full_df_I84_east_time_sub_2week_day_PP_90 %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_sub_2week_day_PP_90)
```

# Inpute missing values with observation at adjacent stations
```{r}
full_df_I84_east_station_sub <- full_df_I84_east_station %>%
                                filter(is.na(speed), !(time_id %in% speed_NA_df$time_id)) %>%
                                mutate(speed_new=ifelse(!is.na(speed_slag1), speed_slag1, speed_slead1),
                                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag2), speed_slag2, speed_new),
                                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead2), speed_slead2, speed_new),
                                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag3), speed_slag3, speed_new),
                                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead3), speed_slead3, speed_new))

# 0.0978 percent missing values are not inputed  
nrow(full_df_I84_east_station_sub %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_station_sub)
```

## Input missing value with earlier/later and adjacent stationid observations
90 minute range & previous/later days & the same day of week of previous or next 2 week & adjacent stations 
```{r}
# Input missing value with earlier/later and adjacent stationid observations ====
full_df_I84_east_time_station <- full_df_I84_east_time %>%
                left_join(full_df_I84_east_station)  %>%
                arrange(date, time_hour, time_min5, stationid) %>%
                filter(is.na(speed), !(time_id %in% speed_NA_df$time_id)) %>%
                mutate(speed_new=ifelse(!is.na(speed_tlag1), speed_tlag1, NA),
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg), speed_avg, speed_new), # average speed
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag1), speed_avg_lag1, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead1), speed_avg_lead1, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag2), speed_avg_lag2, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead2), speed_avg_lead2, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag3), speed_avg_lag3, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead3), speed_avg_lead3, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag4), speed_avg_lag4, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead4), speed_avg_lead4, speed_new),
                        
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag10), speed_avg_lag10, speed_new),
                        speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead10), speed_avg_lead10, speed_new),
                       
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag1), speed_slag1, speed_new),
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead1), speed_slead1, speed_new),
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag2), speed_slag2, speed_new),
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead2), speed_slead2, speed_new),
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag3), speed_slag3, speed_new),
                       speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead3), speed_slead3, speed_new)) %>%
                 select(date, time_hour, time_min5, stationid, speed, speed_new)     

# 0.0288 percent missing value are not inputed 
nrow(full_df_I84_east_time_station %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_time_station)
```

# Convert station-level data to route-level data 
```{r}
full_df_I84_east_update <- full_df_I84_east_time %>%
                          left_join(full_df_I84_east_station)  %>%
                          arrange(date, time_hour, time_min5, stationid) %>%
                          mutate(day_week=weekdays(date))  %>%
                          filter(!(time_id %in% speed_NA_df$time_id), !(day_week %in% c("Saturday", "Sunday")))  %>%
                          mutate(speed_new=ifelse(is.na(speed), speed_tlag1, speed),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag1), speed_tlag1, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead1), speed_tlead1, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag2), speed_tlag2, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead2), speed_tlead2, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag3), speed_tlag3, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead3), speed_tlead3, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag4), speed_tlag4, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead4), speed_tlead4, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag5), speed_tlag5, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead5), speed_tlead5, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag6), speed_tlag6, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead6), speed_tlead6, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag7), speed_tlag7, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead7), speed_tlead7, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag8), speed_tlag8, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead8), speed_tlead8, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlag9), speed_tlag9, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_tlead9), speed_tlead9, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg), speed_avg, speed_new), # average speed
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag1), speed_avg_lag1, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead1), speed_avg_lead1, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag2), speed_avg_lag2, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead2), speed_avg_lead2, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag3), speed_avg_lag3, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead3), speed_avg_lead3, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag4), speed_avg_lag4, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead4), speed_avg_lead4, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lag10), speed_avg_lag10, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_avg_lead10), speed_avg_lead10, speed_new),
                                 
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag1), speed_slag1, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead1), speed_slead1, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag2), speed_slag2, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead2), speed_slead2, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_slag3), speed_slag3, speed_new),
                                 speed_new=ifelse(is.na(speed_new) & !is.na(speed_slead3), speed_slead3, speed_new)) %>%
                          select(date, time_hour, time_min5, stationid, speed, speed_new)    

nrow(full_df_I84_east_update %>% filter(is.na(speed_new)))/nrow(full_df_I84_east_update)

full_df_I84_east_update_route <- full_df_I84_east_update %>%
                                 na.omit() %>%
                                 mutate(travel_time=ifelse(stationid==1055, (0.81/2)/speed_new, NA),
                                        travel_time=ifelse(stationid==1056, (0.81/2 + 1.27/2)/speed_new, travel_time),
                                        travel_time=ifelse(stationid==1057, (1.27/2 + 1.09/2)/speed_new, travel_time),
                                        travel_time=ifelse(stationid==1127, (1.09/2)/speed_new, travel_time)) %>%
                                 group_by(date, time_hour, time_min5) %>%
                                 summarise(speed_seg=(3.69-0.52)/sum(travel_time))
```

# Check low correlation for station lag/lead 
```{r}
full_df_I84_east_station <- full_df_I84_east %>%
                            select(stationid, date, time_hour, time_min5, starttime, AM_PM, speed) %>%
                            arrange(date, time_hour, time_min5, stationid)  %>%
                            group_by(date, AM_PM)  %>%
                            mutate(speed_slag1=lag(speed),
                                   speed_slag2=lag(speed, 5),
                                   speed_slag3=lag(speed, 9),
                                   speed_slead1=lead(speed),
                                   speed_slead2=lead(speed, 5),
                                   speed_slead3=lead(speed, 9)) %>%
                            mutate(time_id=paste(date, time_hour, time_min5, sep="_"),
                                   speed_slag1=ifelse(stationid==1055, NA, speed_slag1),
                                   speed_slag2=ifelse(stationid==1055, NA, speed_slag2),
                                   speed_slag3=ifelse(stationid==1055, NA, speed_slag3),
                                   speed_slead1=ifelse(stationid==1127, NA, speed_slead1),
                                   speed_slead2=ifelse(stationid==1127, NA, speed_slead2),
                                   speed_slead3=ifelse(stationid==1127, NA, speed_slead3)
                            ) %>%
                            as.data.frame()

cor(full_df_I84_east_station$speed, 
    full_df_I84_east_station[, c("speed_slag1", "speed_slag2", "speed_slag3")], 
    use = "complete.obs")

cor(full_df_I84_east_station$speed, 
    full_df_I84_east_station[, c("speed_slead1", "speed_slead2", "speed_slead3")], 
    use = "complete.obs")
```

```{r}
full_df_I84_east_station <- full_df_I84_east %>%
                            select(stationid, date, time_hour, time_min5, starttime, AM_PM, speed) %>%
                            arrange(date, time_hour, time_min5, stationid)  %>%
                            group_by(date, AM_PM)  %>%
                            mutate(speed_slag1=lag(speed),
                                   speed_slag2=lag(speed, 5),
                                   speed_slag3=lag(speed, 9),
                                   speed_slead1=lead(speed),
                                   speed_slead2=lead(speed, 5),
                                   speed_slead3=lead(speed, 9)) %>%
                          #  mutate(time_id=paste(date, time_hour, time_min5, sep="_"),
                          #         speed_slag1=ifelse(stationid==1055, NA, speed_slag1),
                          #         speed_slag2=ifelse(stationid==1055, NA, speed_slag2),
                          #         speed_slag3=ifelse(stationid==1055, NA, speed_slag3),
                          #         speed_slead1=ifelse(stationid==1127, NA, speed_slead1),
                          #         speed_slead2=ifelse(stationid==1127, NA, speed_slead2),
                          #         speed_slead3=ifelse(stationid==1127, NA, speed_slead3)
                          #  ) %>%
                            as.data.frame()

cor(full_df_I84_east_station$speed, 
    full_df_I84_east_station[, c("speed_slag1", "speed_slag2", "speed_slag3")], 
    use = "complete.obs")

cor(full_df_I84_east_station$speed, 
    full_df_I84_east_station[, c("speed_slead1", "speed_slead2", "speed_slead3")], 
    use = "complete.obs")
```

# Others

## Try MICE package 
Reference: https://www.youtube.com/watch?v=An7nPLJ0fsg
Reference: https://datascienceplus.com/imputing-missing-data-with-r-mice-package/

