---
title: "Green Line transit ridership"
date: "11/21/2019"
output: html_document
---

# Setting 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
options(scipen=100)
options(digits=3)

if (!require("pacman")) {install.packages("pacman"); library(pacman)}
p_load(dplyr, lubridate, ggplot2, stargazer, stringr, foreign, scales, ggrepel)
```

# Corridor (Green line from main station to Clackamas center) 
## Load and oranize transit ridership and stop locations data 
```{r}
# Load ridership data 
load("output/intermediate/ridership_w.RData")

# Load GIS bus stop data of con and exp corridors 
gl_exp_busstops <- read.dbf("data/GIS/Green Line/Corridor2/GL_segment_main2clackamas_0.5m_buffer_busstops.dbf", as.is=TRUE) 
gl_con_busstops <- read.dbf("data/GIS/Green Line/Corridor2/I-5_segment_0.5m_buffer_busstops.dbf", as.is=TRUE) 

# 194, 195, 196 are streetcar; no streetcar ridership in ridership dataset
gl_exp_busstops <- gl_exp_busstops %>%
                            filter(!(RTE %in% c("194", "195", "196"))) %>%
                            mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))

gl_con_busstops <- gl_con_busstops %>%
                       filter(!(RTE %in% c("194", "195", "196"))) %>%
                       mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))

gl_ridership_w_exp <- ridership_w %>%
                      filter(rt_dir_stopdesc %in% gl_exp_busstops$rt_dir_stopdesc)

gl_ridership_w_con <- ridership_w %>%
                      filter(rt_dir_stopdesc %in% gl_con_busstops$rt_dir_stopdesc)
```

## Average total weekday boardings at all bus stops and rail stations within the exp and con corridors

### Ridership within the experimental corridor 
```{r}
# Experimental ridership 
gl_ridership_w_exp_summary1 <- gl_ridership_w_exp %>%
                               group_by(ROUTE_NUMBER, PUBLIC_ROUTE_DESCRIPTION, date) %>%
                               summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                               ungroup()

# Select routes operate all the time 
# sort(table(gl_ridership_w_exp_summary1$ROUTE_NUMBER))
# Route 200 is the MAX Green Line
# Routes start service from 2007-03-04 to present 
gl_ridership_w_exp_rt <- gl_ridership_w_exp_summary1 %>%
                          group_by(ROUTE_NUMBER, PUBLIC_ROUTE_DESCRIPTION) %>%
                          summarise(date=min(date), 
                                    freq=n()) %>%
                          arrange(freq) %>%
                          filter(date=="2007-03-04"|ROUTE_NUMBER==200) %>% 
                          as.data.frame()

# Ridership within the experimental corridor by date 
gl_ridership_w_exp_summary2 <- gl_ridership_w_exp %>%
                               filter(ROUTE_NUMBER %in% gl_ridership_w_exp_rt$ROUTE_NUMBER) %>%
                               group_by(date) %>%
                               summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                               ungroup() %>%
                               mutate(Group="Experimental corridor")

# MAX Green Line ridership by date 
gl_ridership_w_exp_summary3 <- gl_ridership_w_exp %>%
                               filter(ROUTE_NUMBER == 200) %>%
                               group_by(date) %>%
                               summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                               ungroup() %>%
                               mutate(Group="MAX Green Line")

# Existing transit service ridership within the experimental corridor 
gl_ridership_w_exp_summary4 <- gl_ridership_w_exp %>%
                               filter(ROUTE_NUMBER %in% gl_ridership_w_exp_rt$ROUTE_NUMBER, ROUTE_NUMBER != 200) %>%
                               group_by(date) %>%
                               summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                               ungroup()  %>%
                               mutate(Group="Existing Transit")

bd_gl_et <- rbind(gl_ridership_w_exp_summary3, gl_ridership_w_exp_summary4)

# Plot of boardings within experimental corridor 
# "Average total weekday boardings at all bus stops and \n rail stations within the experimental corridor"
p_gl_bd_exp <- ggplot(gl_ridership_w_exp_summary2, aes(date, ONES_SUM)) + geom_smooth() + 
                geom_point(size=0.75) + ggtitle("Boardings within the experimental corridor") +
                labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2009-09-12"), linetype='dashed', size=0.3) + 
                theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                      axis.text = element_text(family="Times",size=8),
                      axis.title=element_text(family="Times",size=10),
                      legend.position="bottom",
                      legend.title=element_text(family="Times",size=10, face="bold"),
                      legend.text=element_text(family="Times",size=8)) 
p_gl_bd_exp

# ggsave(file="output/plots/GL plots/Transit/p_gl_bd_exp.png", p_gl_bd_exp, width=5, heigh=3.25)

# Plot MAX Green Line ridership by date 
p_gl_bd_exp_gl <- ggplot(gl_ridership_w_exp_summary3, aes(date, ONES_SUM)) + geom_smooth() +  geom_point(size=0.75) + # ylim (10000, 17000) + 
           ggtitle("Green Line boardings within the experimental corridor") +
           labs(x="Date", y="Boardings") + 
           theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                  axis.text = element_text(family="Times",size=8),
                  axis.title=element_text(family="Times",size=10),
                  legend.position="bottom",
                  legend.title=element_text(family="Times",size=10, face="bold"),
                  legend.text=element_text(family="Times",size=8)) 
p_gl_bd_exp_gl

# ggsave(file="output/plots/GL plots/Transit/p_gl_bd_exp_gl.png", p_gl_bd_exp_gl, width=5, heigh=3.25)  


# Plot existing transit service ridership by date 
p_gl_bd_exp_et <- ggplot(gl_ridership_w_exp_summary4, aes(date, ONES_SUM)) + geom_smooth() +  geom_point(size=0.75)  + # ylim (4000, 7000) + 
           ggtitle("Exisitng transit boardings within the experimental corridor") +
           labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2009-09-12"), linetype='dashed', size=0.3) + 
           theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                  axis.text = element_text(family="Times",size=8),
                  axis.title=element_text(family="Times",size=10),
                  legend.position="bottom",
                  legend.title=element_text(family="Times",size=10, face="bold"),
                  legend.text=element_text(family="Times",size=8)) 
p_gl_bd_exp_et
# ggsave(file="output/plots/p_gl_bd_exp_et.png", p_gl_bd_exp_et, width=5, heigh=3.25) 

# plot of ridership of MAX Orange Line and Exisitng transit data.frame 
p_bd_gl_et <- ggplot(bd_gl_et, aes(date, ONES_SUM, color=Group), group=Group) + geom_smooth(aes(linetype=Group), color="#909497") +
        geom_point(size=0.75) + scale_color_manual(values=c("#A569BD", "#F8C471")) + 
        ggtitle("Boardings at MAX Green Line and exisitng transit \n stops within experimental corridor") +
        labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2009-09-12"), linetype='dashed', size=0.3) + 
        theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
              axis.text = element_text(family="Times",size=8),
              axis.title=element_text(family="Times",size=9),
              legend.position="bottom",
              legend.title=element_text(family="Times",size=9, face="bold"),
              legend.text=element_text(family="Times",size=10))  

p_bd_gl_et
```

### Ridership within the control corridor 
```{r}
# Control ridership 
gl_ridership_w_con_summary1 <- gl_ridership_w_con %>%
                                group_by(ROUTE_NUMBER, PUBLIC_ROUTE_DESCRIPTION, date) %>%
                                summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                                ungroup()

# Select routes that operate all the time 
# sort(table(gl_ridership_w_con_summary1$ROUTE_NUMBER))
# Route 200 is the MAX Green Line
# Routes start service from 2007-03-04 to present 
gl_ridership_w_con_rt <- gl_ridership_w_con_summary1 %>%
                          group_by(ROUTE_NUMBER, PUBLIC_ROUTE_DESCRIPTION) %>%
                          summarise(date=min(date), 
                                    freq=n()) %>%
                          arrange(freq) %>%
                          filter(date=="2007-03-04") %>% #
                          as.data.frame()

gl_ridership_w_con_summary2 <- gl_ridership_w_con %>%
                               filter(ROUTE_NUMBER %in% gl_ridership_w_con_rt$ROUTE_NUMBER) %>%
                               group_by(date) %>%
                               summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                               ungroup() %>%
                               mutate(Group="Control corridor")

# Plot of ridership within control corridor 
# "average total weekday boardings at all bus stops and \n rail stations within the control corridor"
p_gl_bd_con <- ggplot(gl_ridership_w_con_summary2, aes(date, ONES_SUM)) + geom_smooth() + ylim (300, 6000) +
                geom_point(size=0.75) + ggtitle("Boardings within the control corridor") +  
                labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2009-09-12"), linetype='dashed', size=0.3) + 
                theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                      axis.text = element_text(family="Times",size=8),
                      axis.title=element_text(family="Times",size=10),
                      legend.position="bottom",
                      legend.title=element_text(family="Times",size=10, face="bold"),
                      legend.text=element_text(family="Times",size=8)) 

p_gl_bd_con
ggsave(file="output/plots/GL plots/Transit/p_gl_bd_con.png", p_gl_bd_con, width=5, heigh=3.25)
```

### Combine experimental and control corridors 
```{r}
gl_bd_did_df <- rbind(gl_ridership_w_exp_summary2, gl_ridership_w_con_summary2) %>% 
                mutate(Before_after=ifelse(date >= "2009-09-13", 1, 0),
                       Year_1st=ifelse(date>="2009-09-13" & date <= "2010-06-06", 1, 0),
                       Year_2nd=ifelse(date>="2010-09-05" & date <= "2011-06-05", 1, 0),
                       Year_3rd=ifelse(date>="2011-09-04" & date <= "2012-06-03", 1, 0),
                       Year_4th=ifelse(date>="2012-09-02" & date <= "2013-06-02", 1, 0),
                       Year_5th=ifelse(date>="2013-09-01" & date <= "2014-06-01", 1, 0),
                       Year_6th=ifelse(date>="2014-08-31" & date <= "2015-06-07", 1, 0),
                       Year_7th=ifelse(date>="2015-09-13" & date <= "2016-06-05", 1, 0),
                       Year_8th=ifelse(date>="2016-09-04" & date <= "2017-06-04", 1, 0),
                       Year_9th=ifelse(date>="2017-09-03" & date <= "2018-06-03", 1, 0),
                       Year_10th=ifelse(date>="2018-09-02" & date <= "2019-03-03", 1, 0)
                       )

p_gl_bd_ec_cc <- ggplot(gl_bd_did_df, aes(date, ONES_SUM), group=Group) + geom_smooth(aes(linetype=Group), color="#909497") +
                  geom_point(aes(colour=Group), size=0.75) + 
                  ggtitle("Boardings at transit stops within \n the experimental/control corridors") +
                  labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2009-09-12"), linetype='dashed', size=0.3) + 
                  theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                        axis.text = element_text(family="Times",size=8),
                        axis.title=element_text(family="Times",size=10),
                        legend.position="bottom",
                        legend.title=element_text(family="Times",size=10, face="bold"),
                        legend.text=element_text(family="Times",size=10)) #+ ylim(3000, 17000)
p_gl_bd_ec_cc

# ggsave(file="output/plots/p_gl_bd_ec_cc.png", p_gl_bd_ec_cc, width=5, heigh=3.25)
```

### DID regression 
```{r}
# Green Line boarding did base model 
gl_bd_did_base <- lm(ONES_SUM ~ Before_after*Group, data=gl_bd_did_df)

# Multiple time periods 
gl_bd_did_mt <- lm(ONES_SUM ~ Group*Year_1st + Year_2nd*Group + Year_3rd*Group +
                              Year_4th*Group + Year_5th*Group + Year_6th*Group +
                              Year_7th*Group + Year_8th*Group + Year_9th*Group +
                              Year_10th*Group, data=gl_bd_did_df)

stargazer(gl_bd_did_base, gl_bd_did_mt, 
          column.labels = c("Base model", "Multiple time priods"),
          dep.var.labels=c("Boardings"),
          type="text", title="Ridership DID base and multiple time priods models", keep.stat=c("n", "rsq", "adj.rsq", "ser"))
```

#### DID results plots 
```{r}
# Plot for base model 
bm_mx <- matrix(c(1, 1, 1, 1, 1, 1,
                  0, 1, 0, 1, 0, 1,
                  0, 0, 1, 1, 1, 1, 
                  0, 0, 0, 1, 0, 0), nrow=4, byrow=TRUE)

# major arterial speed base models
gl_bd_did_base_coef_mx <- t(as.matrix(summary(gl_bd_did_base)$coefficient[, 1]))


gl_bd_did_base_df <- data.frame(Group=rep(c("Control", "Treated", "Counterfactual"), each=2), 
                             Before_after=rep(c("before", "after"), 3),
                             boardings=as.numeric(gl_bd_did_base_coef_mx %*% bm_mx)) %>%
                     mutate(Before_after=ifelse(Before_after=="after", 1, 0)) 


p_gl_bd_did_base <-  ggplot(gl_bd_did_base_df, aes(x=Before_after, y=boardings, group=Group)) +
                        geom_line(aes(color=Group)) + geom_point() + ylab("Boardings") + 
                        xlab("Time (after=1)")  + scale_x_continuous(breaks=c(0, 1)) + 
                        ggtitle("DID models for ridership base model") + 
                        theme(plot.title = element_text(family="Times",hjust = 0.5, size=12, face="bold"),
                              axis.text = element_text(family="Times",size=8),
                              axis.title=element_text(family="Times",size=12),
                              legend.position="bottom",
                              legend.title=element_text(family="Times",size=12, face="bold"),
                              legend.text=element_text(family="Times",size=8)) 

p_gl_bd_did_base
```


```{r}
# Plot for DID with multiple time periods 
# plot DID models with three year dummy variables 
mt_10yd_mx <- matrix(c(rep(1, 11),
                      rep(1, 11),
                      0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1,
                      
                      0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1),
                      nrow=22, byrow=TRUE)

gl_bd_did_mt_coef_mx <- t(as.matrix(summary(gl_bd_did_mt)$coefficient[, 1]))
gl_bd_did_mt_coef_sesq_mx <- t(as.matrix(summary(gl_bd_did_mt)$coefficient[, 2]))^2


gl_bd_did_mt_df <- data.frame(TreatTime=factor(c("Before", "1st year", "2nd year", "3rd year", "4th year", "5th year", "6th year",
                                              "7th year", "8th year", "9th year", "10th year"), ordered = TRUE, 
                                            levels=c("Before", "1st year", "2nd year", "3rd year", "4th year", "5th year", "6th year",
                                                     "7th year", "8th year", "9th year", "10th year")),
                           Boardings=as.numeric(gl_bd_did_mt_coef_mx %*% mt_10yd_mx),
                           Boardings_se=as.numeric(sqrt(gl_bd_did_mt_coef_sesq_mx %*% mt_10yd_mx))
                           )

# Average treated effect: 10018=sum(summary(gl_bd_did_base)$coefficient[,1])
p_gl_bd_did_mt <-  ggplot(gl_bd_did_mt_df, aes(x=TreatTime, y=Boardings, group=1)) +
        geom_line() + geom_point() + 
        geom_errorbar(aes(ymin=Boardings-1.96*Boardings_se, ymax=Boardings+1.96*Boardings_se, width=.1)) + # , position=position_dodge(0.05)
        geom_hline(yintercept = c(10018), color=c('#999999'), linetype="dashed")+
        xlab("Treatment time periods") + ylab("Boardings") + # ylim(25, 30) +
        ggtitle("Predicted average boardings") + scale_color_manual(values=c('#999999')) +
        theme(plot.title = element_text(family="Times",hjust = 0.5, size=12, face="bold"),
              axis.text = element_text(family="Times",size=8),
              axis.title=element_text(family="Times",size=12),
              legend.position="bottom",
              legend.title=element_text(family="Times",size=12, face="bold"),
              legend.text=element_text(family="Times",size=8)) 

p_gl_bd_did_mt
```
