---
title: "Transit_ridership_analysis"
date: "10/2/2019"
output: html_document
---

# Setting 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
options(scipen=100)
options(digits=3)

if (!require("pacman")) {install.packages("pacman"); library(pacman)}
p_load(dplyr, lubridate, ggplot2, stargazer, stringr, foreign, scales, ggrepel)
```

# Load and organize transit ridership and stop locations data 
```{r}
# # Transit ridership data ====
# # Ridership dataset does not provide streetcar ridership 
# ridership <- read.csv("data/Passenger Census Dump - All.csv", as.is=TRUE)
# 
# month_df <- data.frame(month_num=c(1:12),
#                        month_char=c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC"))
# 
# Weekday ridership 
# ridership_w <- ridership %>%
#                 filter(SERVICE_KEY=="W") %>%
#                 mutate(SUMMARY_BEGIN_DATE=as.character(SUMMARY_BEGIN_DATE),
#                        YEAR=as.numeric(substr(SUMMARY_BEGIN_DATE, 6,9)),
#                        month_char=substr(SUMMARY_BEGIN_DATE, 3,5),
#                        DAY=as.numeric(substr(SUMMARY_BEGIN_DATE, 1,2)),
#                        TIME=substr(SUMMARY_BEGIN_DATE, 11, 18),
#                        rt_dir_stopdesc=paste(ROUTE_NUMBER, DIRECTION, PUBLIC_LOCATION_DESCRIPTION, sep="_"),
#                        hm_char=substr(STOP_TIME, 1, 1),
#                        AM_PM=substr(STOP_TIME, nchar(STOP_TIME)-1, nchar(STOP_TIME)),
#                        AM_Peak=ifelse((hm_char %in% c(6, 7, 8, 9))&AM_PM=="AM", "AM Peak", "Non AM Peak"),
#                        PM_Peak=ifelse((hm_char %in% c(4, 5, 6, 7))&AM_PM=="PM", "PM Peak", "Non PM Peak"),
#                        AM_PM_Peak=ifelse(AM_Peak=="AM Peak"|PM_Peak=="PM Peak", "Peak Periods", "Non Peak Periods")
#                 )  %>%
#                 left_join(month_df) %>%
#                 mutate(date=ymd(paste(YEAR, month_num, DAY, sep="-")))
# 
# save(ridership_w, file = "output/intermediate/ridership_w.RData")

load("~/OrangeGreenLines/output/intermediate/ridership_w.RData")

ridership_w <- ridership_w %>%
               filter(date <= "2018-06-03")

# Use bus stops within corridor (first version of corridor: corridor areas are different in the first and second versions)
# # Load  experimental and control stops GIS bus stop data ====
# control_busstops <-  read.dbf("data/GIS/Control_Corridor_BusStops.dbf", as.is=TRUE)
# experimental_busstops <- read.dbf("data/GIS/Experimental_Corridor_BusStops.dbf", as.is=TRUE)
# 
# # RTE 200 is the MAX Green Line
# control_busstops <- control_busstops %>%
#                     mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))
# 
# # 194, 195, 196 are streetcar; no streetcar ridership in ridership dataset
# # RTE 290 is the MAX Orange Line
# experimental_busstops <- experimental_busstops %>%
#                          filter(!(RTE %in% c("194", "195", "196"))) %>%
#                          mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))
# 
# # extract ridership of control stops from ridership data
# ridership_w_control <- ridership_w %>%
#                           filter(rt_dir_stopdesc %in% control_busstops$rt_dir_stopdesc)
# 
# # all rt_dir_stopdesc are available from ridership dataset
# # sort(unique(ridership_w_rt_control$rt_dir_stopdesc))==sort(unique(control_busstops$rt_dir_stopdesc))
# 
# ridership_w_experimental <- ridership_w %>%
#                                filter(rt_dir_stopdesc %in% experimental_busstops$rt_dir_stopdesc)
# 
# # sort(unique(ridership_w_rt_experimental$rt_dir_stopdesc))==sort(unique(experimental_busstops$rt_dir_stopdesc))
# # There is one stops that do not appear in ridership dataset: SE McLoughlin & Park Ave is not in ridership dataset
# # setdiff(sort(unique(experimental_busstops$rt_dir_stopdesc)), sort(unique(ridership_w_rt_experimental$rt_dir_stopdesc)))

# Use bus stops within corridor1 (second version of corridor)
# Load GIS bus stop data of control corridor and experimental corridor 
control_corridor1_busstops <-  read.dbf("~/OrangeGreenLines/data/GIS/Orange Line/Control_corridor1_busstops.dbf", as.is=TRUE) 
experimental_corridor1_busstops <- read.dbf("~/OrangeGreenLines/data/GIS/Orange Line/Experimental_corridor1_busstops.dbf", as.is=TRUE) 

# RTE 200 is the MAX Green Line
control_corridor1_busstops <- control_corridor1_busstops %>%
                    mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))

# 194, 195, 196 are streetcar; no streetcar ridership in ridership dataset
# RTE 290 is the MAX Orange Line
experimental_corridor1_busstops <- experimental_corridor1_busstops %>%
                          filter(!(RTE %in% c("194", "195", "196"))) %>%
                          mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))

# extract ridership of control stops from ridership data 
ridership_w_rt_control_corridor1 <- ridership_w %>%
                                      filter(rt_dir_stopdesc %in% control_corridor1_busstops$rt_dir_stopdesc)

# all rt_dir_stopdesc are available from ridership dataset
# sort(unique(ridership_w_rt_control_corridor1$rt_dir_stopdesc))==sort(unique(control_corridor1_busstops$rt_dir_stopdesc))
 
ridership_w_rt_experimental_corridorl <- ridership_w %>%
                                          filter(rt_dir_stopdesc %in% experimental_corridor1_busstops$rt_dir_stopdesc)

# There are three stops do not show in riderhship dataset:  "291_0_SE McLoughlin & Park Ave", "33_1_SE Jefferson & Main", "34_1_SE Jefferson & Main" 
# setdiff(sort(unique(experimental_corridor1_busstops$rt_dir_stopdesc)), sort(unique(ridership_w_rt_experimental_corridorl$rt_dir_stopdesc)))

# Avoid recoding 
# Use data from Control_corridor1_busstops.dbf & Experimental_corridor1_busstops.dbf
ridership_w_control <- ridership_w_rt_control_corridor1
ridership_w_experimental <- ridership_w_rt_experimental_corridorl
```


# Transit ridership comparisons 
There are four comparisons. 

## average total weekday boardings at all bus stops and rail stations within the experimental and control corridors
The ridership of the MAX Orange Line accouts for the most of the  increase in ridership. 

```{r}
# Route 2  provides ridership from 2018-09-02 to present
# 291-Orange Night Bus provides ridership from 2015-09-13 to present 
# 290 MAX Orange Line provides ridership from 2015-09-13 to present 
ridership_w_experimental_summary1 <- ridership_w_experimental %>%
                                      filter(date >= "2012-09-02") %>% 
                                      group_by(ROUTE_NUMBER, date) %>%
                                      summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                                      ungroup()

ridership_w_experimental_summary2 <- ridership_w_experimental %>%
                                      filter(!(ROUTE_NUMBER %in% c(2, 291)), date >= "2012-09-02") %>% 
                                      group_by(date) %>%
                                      summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                                      ungroup()

# Plot of ridership within experimental corridor 
# "Average total weekday boardings at all bus stops and \n rail stations within the experimental corridor"
p_bd_ec <- ggplot(ridership_w_experimental_summary2, aes(date, ONES_SUM)) + geom_smooth() + 
           geom_point(size=0.75) + ggtitle("Boardings within the experimental corridor") +
           labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
           theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                 axis.text = element_text(family="Times",size=8),
                 axis.title=element_text(family="Times",size=10),
                 legend.position="bottom",
                 legend.title=element_text(family="Times",size=10, face="bold"),
                 legend.text=element_text(family="Times",size=8)) 


# ggsave(file="~/OrangeGreenLines/output/plots/p_bd_ec.png", p_bd_ec, width=5, heigh=3.25)

# Orange Line ridership 
ridership_w_experimental_summary3 <- ridership_w_experimental %>%
                                      filter(ROUTE_NUMBER %in% c(290)) %>% # Route 290 MAX Orange Line  
                                      group_by(date) %>%
                                      summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                                      ungroup() 

# The ridership of MAX Orange Line is large 
# plot of boardings of MAX Orange Line   
# "Average total weekday boardings at MAX Orange Line \n stations within the experimental corridor"
p_bd_ol <- ggplot(ridership_w_experimental_summary3, aes(date, ONES_SUM)) + geom_smooth() +  geom_point(size=0.75) + ylim (4000, 7000) + 
           ggtitle("Orange Line boardings within the experimental corridor") +
           labs(x="Date", y="Boardings") + 
           theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                  axis.text = element_text(family="Times",size=8),
                  axis.title=element_text(family="Times",size=10),
                  legend.position="bottom",
                  legend.title=element_text(family="Times",size=10, face="bold"),
                  legend.text=element_text(family="Times",size=8)) 
p_bd_ol
# ggsave(file="output/plots/p_bd_ol.png", p_bd_ol, width=5, heigh=3.25)  

# Existing bus ridership within experimental corridor 
ridership_w_experimental_summary4 <- ridership_w_experimental %>%
                                      filter(!(ROUTE_NUMBER %in% c(2, 290, 291)), date >= "2012-09-02") %>% 
                                      group_by(date) %>%
                                      summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
                                      ungroup()


p_bd_ec_no_ol <- ggplot(ridership_w_experimental_summary4, aes(date, ONES_SUM)) + geom_smooth() +  geom_point(size=0.75)  + ylim (4000, 7000) + 
           ggtitle("Bus boardings within the experimental corridor") +
           labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
           theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                  axis.text = element_text(family="Times",size=8),
                  axis.title=element_text(family="Times",size=10),
                  legend.position="bottom",
                  legend.title=element_text(family="Times",size=10, face="bold"),
                  legend.text=element_text(family="Times",size=8)) 
p_bd_ec_no_ol
# ggsave(file="output/plots/p_bd_ec_no_ol.png", p_bd_ec_no_ol, width=5, heigh=3.25) 


# existing trasnit ridership by route 
ridership_w_experimental_summary5 <- ridership_w_experimental %>%
        filter(!(ROUTE_NUMBER %in% c(2, 290, 291)), date >= "2012-09-02") %>% 
        group_by(ROUTE_NUMBER, date) %>%
        summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
        ungroup() %>%
        mutate(ROUTE_NUMBER=as.factor(ROUTE_NUMBER)) %>%
        mutate(label = if_else(date == max(date), as.character(ROUTE_NUMBER), NA_character_)) %>%
        as.data.frame()

# Plot of boardings of existing bus within experimental corridor by route 
p_bd_ec_et_rt <- ggplot(ridership_w_experimental_summary5, aes(date, ONES_SUM, color=ROUTE_NUMBER), group=ROUTE_NUMBER) + # geom_smooth(aes(linetype=ROUTE_NUMBER), color="#909497") +
          geom_line(size=0.3) + # scale_color_manual(values=c("#A569BD", "#F8C471")) + 
          ggtitle("Boardings at exisitng transit \n stops within experimental corridor") +
          labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
          #  scale_x_date(breaks="month", labels=date_format("%b")) +
          scale_x_date(breaks = pretty_breaks(20))  + 
          theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                axis.text = element_text(family="Times",size=8),
                axis.title=element_text(family="Times",size=9),
                legend.position = "none"
                # legend.position="bottom",
                # legend.title=element_text(family="Times",size=9, face="bold"),
                # legend.text=element_text(family="Times",size=10)
                ) + geom_label_repel(aes(label = label),
                           nudge_x = 1,
                           na.rm = TRUE)

p_bd_ec_et_rt

# ggsave(file="output/plots/p_bd_ec_et_rt.png", p_bd_ec_et_rt, width=5, heigh=3.25) 

# ridership of Orange Line and Existing transit data.frame 
ridership_w_experimental_summary3 <- ridership_w_experimental_summary3 %>%
                                     mutate(Group="MAX Orange Line")

ridership_w_experimental_summary4 <- ridership_w_experimental_summary4 %>%
                                     mutate(Group="Existing Transit")

rs_ol_et <- rbind(ridership_w_experimental_summary3, ridership_w_experimental_summary4)

# plot of ridership of MAX Orange Line and Exisitng transit data.frame 
p_rs_ol_et <- ggplot(rs_ol_et, aes(date, ONES_SUM, color=Group), group=Group) + geom_smooth(aes(linetype=Group), color="#909497") +
  geom_point(size=0.75) + scale_color_manual(values=c("#A569BD", "#F8C471")) + 
  ggtitle("Boardings at MAX Orange Line and exisitng transit \n stops within experimental corridor") +
  labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + ylim(4000, 7000) +
  theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
        axis.text = element_text(family="Times",size=8),
        axis.title=element_text(family="Times",size=9),
        legend.position="bottom",
        legend.title=element_text(family="Times",size=9, face="bold"),
        legend.text=element_text(family="Times",size=10))  

p_rs_ol_et

# ggsave(file="p_rs_ol_et.png", p_rs_ol_et, width=5, heigh=3.25)
# ggsave(file="output/plots/p_rs_ol_et.png", p_rs_ol_et, width=5, heigh=3.25)

```


Transit ridership decreased within the control corridor after the operation of the MAX Orange Line. 
```{r}
# Route 2 only provide ridership from 2018-09-02 to present
# Route 73 only provide ridership from 2016-09-04 to present 
# Route 74 only provide ridership form 2018-03-04
# Ridership of RTE 200 MAX Green Line is from 2012-09-02 to present 
ridership_w_control_summary1 <- ridership_w_control %>%
  filter(date >= "2012-09-02") %>%
  group_by(ROUTE_NUMBER, date) %>%
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() 

ridership_w_control_summary2 <- ridership_w_control %>%
  filter(!(ROUTE_NUMBER %in% c(2, 73, 74)), date >= "2012-09-02") %>%
  group_by(date) %>%
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() 

# Plot of ridership within control corridor 
# "average total weekday boardings at all bus stops and \n rail stations within the control corridor"
p_bd_cc <- ggplot(ridership_w_control_summary2, aes(date, ONES_SUM)) + geom_smooth() + ylim (11000, 16000) +
           geom_point(size=0.75) + ggtitle("Boardings within the control corridor") +  
           labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
           theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                axis.text = element_text(family="Times",size=8),
                axis.title=element_text(family="Times",size=10),
                legend.position="bottom",
                legend.title=element_text(family="Times",size=10, face="bold"),
                legend.text=element_text(family="Times",size=8)) 

p_bd_cc
# ggsave(file="output/plots/p_bd_cc.png", p_bd_cc, width=5, heigh=3.25)
```

```{r}
# Orangeize ridership dataframe for DID regression 
ridership_w_experimental_summary2 <- ridership_w_experimental_summary2 %>%
  mutate(Group="Experimental corridor")

ridership_w_control_summary2 <- ridership_w_control_summary2 %>%
  mutate(Group="Control corridor")

rs_did_df <- rbind(ridership_w_experimental_summary2, ridership_w_control_summary2) %>% 
             mutate(Before_after=ifelse(date >= "2015-09-13", 1, 0),
                    Year_1st=ifelse(date>="2015-09-13" & date <= "2016-06-05", 1, 0),
                    Year_2nd=ifelse(date>="2016-09-04" & date <= "2017-06-04", 1, 0),
                    Year_3rd=ifelse(date>="2017-09-03" & date <= "2018-06-03", 1, 0))

p_bd_ec_cc <- ggplot(rs_did_df, aes(date, ONES_SUM), group=Group) + geom_smooth(aes(linetype=Group), color="#909497") +
  geom_point(aes(colour=Group), size=0.75) + 
  ggtitle("Boardings at transit stops within \n the experimental/control corridors") +
  labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
  theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
        axis.text = element_text(family="Times",size=8),
        axis.title=element_text(family="Times",size=10),
        legend.position="bottom",
        legend.title=element_text(family="Times",size=10, face="bold"),
        legend.text=element_text(family="Times",size=10)) + ylim(3000, 17000)
p_bd_ec_cc

# ggsave(file="output/plots/p_bd_ec_cc.png", p_bd_ec_cc, width=5, heigh=3.25)
```

### DID regression analysis 
```{r}
# Base model 
rs_did_base <- lm(ONES_SUM ~ Before_after*Group, data=rs_did_df)

# Multiple time periods 
rs_did_mt <- lm(ONES_SUM ~ Group*Year_1st + Group*Year_2nd + Group*Year_3rd, data=rs_did_df)

stargazer(rs_did_base, rs_did_mt, 
          column.labels = c("Base model", "Multiple time priods"),
          dep.var.labels=c("Boardings"),
          type="text", title="Ridership base and multiple time priods", keep.stat=c("n", "rsq", "adj.rsq", "ser"))
```

### Plots for results 
```{r}
# Plot for base model 
bm_mx <- matrix(c(1, 1, 1, 1, 1, 1,
                  0, 1, 0, 1, 0, 1,
                  0, 0, 1, 1, 1, 1, 
                  0, 0, 0, 1, 0, 0), nrow=4, byrow=TRUE)

# major arterial speed base models
rs_did_base_coef_mx <- t(as.matrix(summary(rs_did_base)$coefficient[, 1]))


rs_did_base_df <- data.frame(Group=rep(c("Control", "Treated", "Counterfactual"), each=2), 
                             Before_after=rep(c("before", "after"), 3),
                             boardings=as.numeric(rs_did_base_coef_mx %*% bm_mx)) %>%
                  mutate(Before_after=ifelse(Before_after=="after", 1, 0)) 

p_rs_did_base <-  ggplot(rs_did_base_df, aes(x=Before_after, y=boardings, group=Group)) +
                    geom_line(aes(color=Group)) + geom_point() + ylab("Boardings") + 
                    xlab("Time (after=1)")  + scale_x_continuous(breaks=c(0, 1)) + 
                    ggtitle("DID models for ridership") + 
                    theme(plot.title = element_text(family="Times",hjust = 0.5, size=12, face="bold"),
                          axis.text = element_text(family="Times",size=8),
                          axis.title=element_text(family="Times",size=12),
                          legend.position="bottom",
                          legend.title=element_text(family="Times",size=12, face="bold"),
                          legend.text=element_text(family="Times",size=8)) 

p_rs_did_base

# ggsave(file="output/plots/DID plots/p_rs_did_base.png", p_rs_did_base, width=5, heigh=4)


# plot DID models with three year dummy variables 
mt_3yd_mx <- matrix(c(rep(1, 4),
                      rep(1, 4),
                      0, 1, 0, 0, 
                      0, 0, 1, 0,
                      0, 0, 0, 1,
                      0, 1, 0, 0, 
                      0, 0, 1, 0,
                      0, 0, 0, 1), nrow=8, byrow=TRUE)

# AM outbound 
rs_did_mt_coef_mx <- t(as.matrix(summary(rs_did_mt)$coefficient[, 1]))
rs_did_mt_coef_sesq_mx <- t(as.matrix(summary(rs_did_mt)$coefficient[, 2]))^2

rs_did_mt_df <- data.frame(TreatTime=factor(c("Before", "1st year", "2nd year", "3rd year"), ordered = TRUE, 
                                                     levels=c("Before", "1st year", "2nd year", "3rd year")),
                                    Boardings=as.numeric(rs_did_mt_coef_mx %*% mt_3yd_mx),
                                    Boardings_se=as.numeric(sqrt(rs_did_mt_coef_sesq_mx %*% mt_3yd_mx)))

# Average treated effect: 11744=sum(summary(rs_did_base)$coefficient[,1])
p_rs_did_mt <-  ggplot(rs_did_mt_df, aes(x=TreatTime, y=Boardings, group=1)) +
  geom_line() + geom_point() + 
  geom_errorbar(aes(ymin=Boardings-1.96*Boardings_se, ymax=Boardings+1.96*Boardings_se, width=.1)) + # , position=position_dodge(0.05)
  geom_hline(yintercept = c(11744), color=c('#999999'), linetype="dashed")+
  xlab("Treatment time periods") + ylab("Boardings") + # ylim(25, 30) +
  ggtitle("Predicted average boardings") + scale_color_manual(values=c('#999999')) +
  theme(plot.title = element_text(family="Times",hjust = 0.5, size=12, face="bold"),
        axis.text = element_text(family="Times",size=8),
        axis.title=element_text(family="Times",size=12),
        legend.position="bottom",
        legend.title=element_text(family="Times",size=12, face="bold"),
        legend.text=element_text(family="Times",size=8)) 

p_rs_did_mt

# ggsave(file="output/plots/DID plots/p_rs_did_mt.png", p_rs_did_mt, width=5, heigh=4)
```

## compared changes in average total weekday ridership of east–west transit lines within the study area
These are lines that are alternatives to the Expo Line within the experimental corridor. Ridership is the total boardings along entire routes both directions).

Experimental: Routes 19, 30, 70, 291 cross the screenline  
Control: East-west route 9, 17 cross the screenliens 
```{r}
# Routes 
control_RT <- sort(unique(ridership_w_control_summary1$ROUTE_NUMBER))
experimental_RT <- sort(unique(ridership_w_experimental_summary1$ROUTE_NUMBER))

# Experimental corridor 
# East-west routes: 9
# North-south routes: 30, 32, 33, 34, 70, 75, 99, 291
# Route 17: half north-south and half east-west
# Route 19: 2/3 east-west and 1/3 north-south 
# Routh 29: short length 
# Route 66 west part/short 
# Route 99 short east west; 
# Route 290 MAX Orange Line

# Routes 19, 30, 70, 290, 291 cross the screenline  
ridership_experimental  <- ridership_w %>% 
  filter(ROUTE_NUMBER %in% c(19, 30, 70, 290, 291), date >= "2012-09-02")  %>% 
  group_by(date)  %>% 
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() 

ggplot(ridership_experimental, aes(date, ONES_SUM)) + geom_point() + geom_smooth()  +
  ggtitle("average total weekday ridership of east–west \n transit lines within the experimental corridor (include Orange Line")


# Routes 19, 30, 70, 291 cross the screenline  
ridership_experimental2  <- ridership_w %>% 
  filter(ROUTE_NUMBER %in% c(19, 30, 70, 291), date >= "2012-09-02")  %>% 
  group_by(date)  %>% 
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() 

ggplot(ridership_experimental2, aes(date, ONES_SUM)) + geom_point() + geom_smooth()  +
  ggtitle("average total weekday ridership of east–west \n transit lines within the experimental corridor (exclude Orange Line")

# Control corridor  
# East-west routes: 9, 10, 14, 17, 
# North-south: 71, 72, 75, 
# Route 66 west part/short 
# 200 MAX Green Line 

# East-west route 9, 17 cross the screenliens 
ridership_control  <- ridership_w %>% 
  filter(ROUTE_NUMBER %in% c(9, 17), date >= "2012-09-02")  %>% 
  group_by(date)  %>% 
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() 

ggplot(ridership_control, aes(date, ONES_SUM)) + geom_point() + geom_smooth() + ylim (11000, 20000) +
  ggtitle("average total weekday ridership of east–west \n transit lines within the control corridor")

```

##  Weekday peak-period (AM and PM peaks combined) bidirectional person throughput by east–west Metro bus and rail lines across the same screenlines, before and after.



```{r}
# Ridership of parallel routes during peak 
ridership_rt_experimental_peak  <-  ridership_w %>% 
  filter(ROUTE_NUMBER %in% c(19, 30, 70, 290, 291), date >= "2012-09-02", AM_PM_Peak=="Peak Periods")  %>% 
  group_by(date)  %>% 
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(Group="Experimental corridor")

ggplot(ridership_rt_experimental_peak, aes(date, ONES_SUM)) + geom_point() + geom_smooth() + 
  ggtitle("Weekday peak-period (AM and PM) bidirectional person \n throughput by east–west transit lines across the experimental corridor")

# Control ridership for peak period    
ridership_rt_control_peak  <- ridership_w %>% 
  filter(ROUTE_NUMBER %in% c(9, 17), date >= "2012-09-02", AM_PM_Peak=="Peak Periods")  %>% 
  group_by(date)  %>% 
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(Group="Control corridor")

ggplot(ridership_rt_control_peak, aes(date, ONES_SUM)) + geom_point() + geom_smooth() + 
  ggtitle("Weekday peak-period (AM and PM) bidirectional person \n throughput by east–west transit lines across the control corridor")


ridership_rt_ec_cc_peak <- rbind(ridership_rt_experimental_peak, ridership_rt_control_peak)

p_rs_rt_ec_cc <- ggplot(ridership_rt_ec_cc_peak, aes(date, ONES_SUM), group=Group) + geom_smooth(aes(linetype=Group), color="#909497") +
                  geom_point(aes(colour=Group), size=0.75) + 
                  ggtitle("Peak-periods bidirectional boardings of transit \n routes across control/experimental corridors") +
                  labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
                  theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                        axis.text = element_text(family="Times",size=8),
                        axis.title=element_text(family="Times",size=10),
                        legend.position="bottom",
                        legend.title=element_text(family="Times",size=10, face="bold"),
                        legend.text=element_text(family="Times",size=10)) # + ylim(3000, 17000)

p_rs_rt_ec_cc

# ggsave(file="output/plots/p_rs_rt_ec_cc.png", p_rs_rt_ec_cc, width=5, heigh=3.25)


ridership_rt_experimental_peak2  <-  ridership_w %>% 
  filter(ROUTE_NUMBER %in% c(19, 30, 70, 290, 291), date >= "2012-09-02", AM_PM_Peak=="Peak Periods")  %>% 
  mutate(Transit=ifelse(ROUTE_NUMBER==290, "MAX Orange Line", "Existing parallel bus lines"))  %>% 
  group_by(Transit, date)  %>% 
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE)) %>%
  ungroup() 

 ggplot(ridership_rt_experimental_peak2, aes(date, ONES_SUM, color=Transit), group=Transit) +       geom_smooth(aes(linetype=Transit), color="#909497") +
                  geom_point(size=0.75) + scale_color_manual(values=c("#A569BD", "#F8C471")) + 
                  ggtitle("Peak periods bidirectional boardings of MAX \n Orange Line and traverse bus lines") +
                  labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
                  theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                        axis.text = element_text(family="Times",size=8),
                        axis.title=element_text(family="Times",size=9),
                        legend.position="bottom",
                        legend.title=element_text(family="Times",size=9, face="bold"),
                        legend.text=element_text(family="Times",size=10))  #+ ylim(4000, 14000) 


# ggsave(file="output/plots/p_rs_rt_ol_et.png", p_rs_rt_ol_et, width=5, heigh=3.25)
```


## Boardings and alightings at specific bus stops within a quarter-mile walking radius of MAX Orange Line stations.
```{r}
bus_stops_quarter_mile <-  read.dbf("~/OrangeGreenLines/data/GIS/Orange Line/Bus_stops_quarter_mile_withoutMAX_SC.dbf", as.is=TRUE) 

bus_stops_quarter_mile <- bus_stops_quarter_mile %>%
  mutate(rt_dir_stopdesc=paste(RTE, DIR, LOCATION, sep="_"))  %>%
  select(RTE, DIR, LOCATION,  rt_dir_stopdesc) # , join_STATI

ridership_bus_stops_quarter_mile <- ridership_w %>%
  filter(rt_dir_stopdesc %in% bus_stops_quarter_mile$rt_dir_stopdesc) %>%
  left_join(bus_stops_quarter_mile)  


ridership_bus_stops_quarter_mile_summary1 <- ridership_bus_stops_quarter_mile %>%
  filter(!(ROUTE_NUMBER %in% c(2)), date >= "2012-09-02") %>%
  group_by(date) %>%
  summarise(ONES_SUM=sum(ONS, na.rm = TRUE),
            OFFS_SUM=sum(OFFS, na.rm = TRUE))

# plot of boardings at specific bus stops within q quarter-mile working radius of MAX Orange Line stations 
p_bd_bs_qm  <-  ggplot(ridership_bus_stops_quarter_mile_summary1, aes(date, ONES_SUM)) + geom_smooth() + 
                geom_point(size=0.75) + ggtitle("Boardings at specific bus stops within a quarter-mile \n walking radius of MAX Orange Line stations") +
                labs(x="Date", y="Boardings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
                theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                      axis.text = element_text(family="Times",size=8),
                      axis.title=element_text(family="Times",size=10),
                      legend.position="bottom",
                      legend.title=element_text(family="Times",size=10, face="bold"),
                      legend.text=element_text(family="Times",size=8)) 
p_bd_bs_qm  
  
# ggsave(file="output/plots/p_bd_bs_qm.png", p_bd_bs_qm, width=5, heigh=3.25)  



# plot of alightings at specific bus stops within q quarter-mile working radius of MAX Orange Line stations   
p_al_bs_qm<- ggplot(ridership_bus_stops_quarter_mile_summary1, aes(date, OFFS_SUM)) + geom_smooth() + 
             geom_point(size=0.75) +ggtitle("Alightings at specific bus stops within a quarter-mile \n walking radius of MAX Orange Line stations") +
             labs(x="Date", y="Alightings") + geom_vline(xintercept=ymd("2015-09-12"), linetype='dashed', size=0.3) + 
             theme(plot.title = element_text(family="Times",hjust = 0.5, size=10, face="bold"),
                    axis.text = element_text(family="Times",size=8),
                    axis.title=element_text(family="Times",size=10),
                    legend.position="bottom",
                    legend.title=element_text(family="Times",size=10, face="bold"),
                    legend.text=element_text(family="Times",size=8)) 

p_al_bs_qm

# ggsave(file="output/plots/p_al_bs_qm.png", p_al_bs_qm, width=5, heigh=3.25)  


```

